A user-extensible and adaptable parser architecture

### 2008-05-30 papers/sgai-2008/paper.tex
Small, simple, not too intrusive changes to shorten the paper. Still needs lots more shortening.

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that user have an in-depth
    knowledge and understanding of the parser's internal workings.  This
    paper presents a novel parsing architecture which aims to make the
    process of parsing new inputs as simple as possible, enabling users to
    trivially add new rules or adapt existing rules (to parse variants of
    existing inputs) and relatively easily add new actions (to process a
    previously unknown class of input).  The architecture scales linearly
    as the number of rules and/or size of input increases, making it
    suitable for parsing large corpora or months of accumulated data.
}

\today{}

XXX DOES THE ABSTRACT NEED TO BE LONGER\@?  IS IT A GOOD TITLE\@?

XXX DUE MONDAY JUNE 2$^{ND}$ 2008\@.  PRESENTATION PAPERS\@: MAX 14
PAGES\@.  POSTER PAPERS\@: MAX 6 PAGES\@.  BLACK AND WHITE ONLY\@.

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing new measures to supplement the
existing filters where there is a deficiency.  The approach chosen to
measure performance was to analyse the log files produced by the mail
server, rather than modifying the mail server to generate performance
statistics; this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the \SMTP{} server in use.  The variety occurs for
several reasons:

@@

    \item [Actions]  The actions perform the work required to deal with a
        category of log lines, e.g.\ creating a new data structure when a
        log line indicating a remote client connected is processed.

    \item [Rules]  The rules are responsible for matching log lines,
        specifying the action to invoke and the data to be extracted from
        the log line.

@@

The framework provides utility and support functions for the actions and
rules.  It provides shared storage to pass data between actions, most
notably the data structures used to represent each connection and mail.  It
loads and validates rules, controls parsing of log files, invokes actions,
tracks how often each rule matches (to optimise rule ordering, see
\sectionref{Rule ordering}), and stores results in the database on request.
The framework is independent of the parser being written: the same
functionality will be required from the framework for most or all parsers.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It links actions and rules, allowing either to be
improved independently of the other.  The framework is the core of the
architecture and is deliberately quite simple: the rules deal with the
varying log lines, and the actions deal with the intricacies and
complications encountered when parsing.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser will need to write the actions it requires from scratch, unless
it's extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{SAVE\_BY\_QUEUEID}, the action
responsible for handling informative inputs: these are log lines that
supplement the data gathered from other log lines (such as rejections), but
that don't by themselves represent interesting events.  The third most
common action is, perhaps surprisingly, \texttt{IGNORE}: there are many log
lines which are not of interest to the parser, e.g.\ warning messages about
DNS mismatches, or log lines lacking a way to identify which data structure
they should be associated with.\footnote{Postfix allocates a queueid
(explained in the glossary,~\sectionref{Glossary}) to each mail; most log
lines contain the queueid of the associated mail, but some log lines
(particularly with older versions of Postfix) lack a queueid, preventing
their association with a particular data structure.}  The parser warns
about every unparsed log line, to alert the user that they need to alter or
extend their rules; the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The remaining actions have only one or two associated rules: in some cases
this is because there will only ever be one type of log line for that
action, e.g.\ all log lines showing that a remote client has connected are
matched by a single rule and handled by the \texttt{CONNECT} action; in
other cases the action is required to address a deficiency in the log
files, or a complication which arises during parsing.

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour or resolve
challenges, with all the attendant risks of adversely affecting the
existing, working parser, when a new requirement arises a new, independent
action can be written to satisfy it.  In some cases the new action will
require the cooperation of other actions, e.g.\ to set or check a flag;
there is a possibility of introducing failure by modifying existing
actions, but modifications will be smaller and occur less frequently than
with a monolithic architecture, and thus failures should be less likely and
easier to test for and diagnose.  The parser can be written in an Object
Oriented~\cite{Wikipedia-object-orientation} style, allowing sub-classes to
extend or override actions in addition to adding new actions; because each
action is an independent function, the sub-class need only modify the
action it is interested in, rather than reproducing large chunks of
functionality.

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.  Some were the result of deficiencies in Postfix's logging,
which were improved or resolved by later versions of Postfix; others were
due to the vagaries of process scheduling, client behaviour, and
administrative actions.  All were successfully accommodated in the parser:
adding new actions was enough for several of the complications; others
required modifications to a single existing action to work around the
difficulties; the remainder were resolved by adapting actions to pass
information to subsequent actions (via the framework), which then check for
the extra information and change their behaviour appropriately.

@@

Rules are responsible for categorising inputs, specifying both the data to
be extracted from each log line and the action that should be invoked.
This implementation stores the rules in the same \SQL{} database that the
results are stored in; this removes any doubt as to which set of rules was
used to produce a set of results.  Other implementations are free to store
the rules in whatever fashion suits their needs.  Each rule defines several
attributes:

@@

    \item [priority] This is the user-configurable companion to hits:
        priority overrides hits when sorting rules.  This allows more
        specific rules to take precedence over more general rules
        (described in \sectionref{overlapping rules}).

@@

Adding new rules is simple: simply define all the required attributes and
append the new rule to wherever the existing rules are stored.  The
software developed for this project includes a utility based on
\SLCT{}~\cite{slct-paper} to aid in producing \regexes{} to parse
previously unparsed inputs.

@@

XXX DO I NEED TO DISCUSS OVERLAPPING RULES\@?

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.\footnote{It may be possible to parse each
rule's regex and determine if any overlap.  The author has not attempted to
do this: given the tremendous power and flexibility of regexes,
particularly Perl's regexes~\cite{perlre}, it would be an extremely
difficult task.} Unintentionally overlapping rules lead to inconsistent
parsing and data extraction because the first matching rule wins and the
order in which rules are tried against each line may change between log
files.  Overlapping rules are frequently a requirement, allowing a more
specific rule to match some log lines and a more general rule to match the
majority, e.g.\ separating \SMTP{} delivery to specific sites from \SMTP{}
delivery to the rest of the world.  The algorithm provides a facility for
specifying the order of overlapping rules: the priority field in each rule
(defaults to zero).  Rules are sorted by priority, highest first, and then
rules with the same priority are sorted by the number of successful matches
when parsing the previous log file.  Negative priorities may be useful for
catchall rules.

@@

Saving results to the database was disabled for the test runs, as that
dominates the run time of the program, and the tests are aimed at measuring
the speed of the parser rather than the speed of the database and the disks
the database is stored on.

\subsection{Architecture Scalability: Input size}

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Figure~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB, and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by fig.~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time, and ratio of number of log lines vs execution time (higher
is better in both cases).  As the reader can see the ratios are quite
tightly banded across the graph, showing that the algorithm scales
linearly: the much larger log files at points 62--68 on the X axis in
fig.~\refwithpage{execution time vs file size vs number of lines graph}
actually cause the ratio to increase (i.e.\ improve), rather than decrease.
The strange behaviour where larger log files are parsed more efficiently is
explained in \sectionref{Why are there dips in the graphs?}.
\showgraph{build/plot-normal-filesize-numlines}{Execution time vs file size
vs number of lines}{execution time vs file size vs number of lines graph}
\showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file size
and number of lines to execution time}{execution time vs file size vs
number lines factor}

>>>>

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  This paper presents a
    novel parsing architecture which aims to make the process of parsing
    new inputs as simple as possible, enabling users to trivially add new
    rules or adapt existing rules (to parse variants of existing inputs)
    and relatively easily add new actions (to process a previously unknown
    class of input).  The architecture scales linearly as the number of
    rules and/or size of input increases, making it suitable for parsing
    large corpora or months of accumulated data.
}

XXX SHOULD THE ABSTRACT BE LONGER\@?  IS IT A GOOD TITLE\@?

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing new measures to supplement the
existing filters where there is a deficiency.  The approach chosen to
measure performance was to analyse the log files produced by the mail
server, rather than modifying the mail server to generate performance
statistics; this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the \SMTP{} server in use.  The variety in log
lines occurs for several reasons:

@@

    \item [Actions]  The actions perform the work required to deal with a
        category of log lines, e.g.\ processing data from rejections.

    \item [Rules]  The rules are responsible for classifying or matching
        log lines, specifying the action to invoke and the data to be
        extracted from the log line.

@@

The framework provides utility and support functions for the actions and
rules.  It provides shared storage to pass data between actions, most
notably the data structures used to represent each connection and mail.  It
loads and validates rules, controls parsing of log files, invokes actions,
tracks how often each rule matches (to optimise rule ordering, see
\sectionref{Rule ordering}), and stores results in the database.  The
framework is independent of the parser being written: the same
functionality will be required for most or all parsers.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  The framework
is the core of the architecture and is deliberately quite simple: the rules
deal with the varying log lines, and the actions deal with the intricacies
and complications encountered when parsing.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser will need to write the actions it requires from scratch, unless
it's extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{SAVE\_BY\_QUEUEID}, the action
responsible for handling informative inputs: these are log lines that
supplement the data gathered from other log lines (such as rejections), but
that don't by themselves represent interesting events.  The third most
common action is, perhaps surprisingly, \texttt{IGNORE}: there are many log
lines which are not of interest to the parser, e.g.\ warning messages about
DNS mismatches, or log lines lacking a way to identify which data structure
they should be associated with.  The parser warns about every unparsed log
line, to alert the user that they need to alter or extend their rules; the
\texttt{IGNORE} action does nothing when executed, allowing uninteresting
log lines to be parsed without causing any effects.  The remaining actions
have only one or two associated rules: in some cases this is because there
will only ever be one type of log line for that action, e.g.\ all log lines
showing that a remote client has connected are matched by a single rule and
handled by the \texttt{CONNECT} action; in other cases the action is
required to address a deficiency in the log files, or a complication which
arises during parsing.

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour or resolve
challenges, with all the attendant risks of adversely affecting the
existing, working parser, when a new requirement arises an independent
action can be written to satisfy it.  In some cases the new action will
require the cooperation of other actions, e.g.\ to set or check a flag;
there is a possibility of introducing failure by modifying existing
actions, but modifications will be smaller and occur less frequently than
with a monolithic architecture, and thus failures should be less likely and
easier to test for and diagnose.  The parser can be written in an object
oriented style, allowing sub-classes to extend or override actions in
addition to adding new actions; because each action is an independent
function, the sub-class need only modify the action it is interested in,
rather than reproducing large chunks of functionality.

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.  Some were the result of deficiencies in Postfix's logging,
(some of those deficiencies were improved or resolved by later versions of
Postfix); others were due to the vagaries of process scheduling, client
behaviour, and administrative actions.  All were successfully accommodated
in the parser: adding new actions was enough for several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
actions to pass information to subsequent actions (via the framework),
those subsequent actions then check for the extra information and change
their behaviour appropriately.

@@

Rules are responsible for categorising inputs, specifying both the data to
be extracted from each log line and the action that should be invoked.  The
Postfix log parser stores the rules in the same \SQL{} database that the
results are stored in; this removes any doubt as to which set of rules was
used to produce a set of results.  Other implementations are free to store
the rules in whatever fashion suits their needs.  Each rule defines several
attributes:

@@

    \item [priority] This is the user-configurable companion to hits:
        priority overrides hits when sorting rules, allowing more
        specific rules to take precedence over more general rules.

@@

Adding new rules is simple: simply define all the required attributes and
append the new rule to wherever the existing rules are stored.  The
software developed for this project includes a utility based on
SLCT~\cite{slct-paper} to aid in producing \regexes{} to parse previously
unparsed inputs.

@@

%XXX DO I NEED TO DISCUSS OVERLAPPING RULES\@?

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \SMTP{} delivery to specific
sites from \SMTP{} delivery to the rest of the world.  The algorithm
provides a facility for specifying the order of overlapping rules: the
priority field in each rule (defaults to zero).  Rules are sorted by
priority, highest first, and then rules with the same priority are sorted
by the number of successful matches when parsing the previous log file.
Negative priorities may be useful for catchall rules.

@@

\subsection{Architecture Scalability: Input Size}

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Figure~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB, and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
the impression that the algorithm scales linearly with input size.
This impression is borne out by fig.~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratios of file size vs
execution time, and number of log lines vs execution time (higher
is better in both cases).  As the reader can see the ratios are quite
tightly banded across the graph, showing that the algorithm scales
linearly: the much larger log files at points 62--68 on the X axis in
fig.~\refwithpage{execution time vs file size vs number of lines graph}
actually cause the ratio to increase (i.e.\ improve), rather than decrease.
The strange behaviour where larger log files are parsed more efficiently is
explained in \sectionref{Why are there dips in the graphs?}.
\showgraph{build/plot-normal-filesize-numlines}{Execution time vs file size
vs number of lines}{execution time vs file size vs number of lines graph}
\showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file size
and number of lines to execution time}{execution time vs file size vs
number lines factor}
